---
title: "tidy-model"
author: "Amanda"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
```

## Introduction 
This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that sound? 
In this report, we examine what factors predict a cricket's chirp rate. 

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab ('Chirp rate (per min.)')
```

## Temperature affects chirp rate 

```{r temp, echo = FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") + 
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping
```{r species, echo=  FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
  
species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```

Based on the scatter plot, both species show a positive relationship, based on the R^2 value 
there is a 98.9% chance that the variance chirp rate is accounted by the temperature.

## Interactions

```{r species histogram, echo = FALSE}
ggplot(crickets, aes(x = rate, fill = species)) +
  geom_histogram(position = 'identity', alpha = 0.7, bins = 8) +
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

```{r interactions, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
species_x_temp_lm <- lm(rate ~ temp + species + temp:species, crickets) 
summary.lm(species_x_temp_lm)

anova(species_lm, species_x_temp_lm)
```

The P-value result of the models show up as 0.25 which is less than 0.05 which means the difference between species-temperature interaction and species-species models are not statistically significant. Therefore, we will only focus on the species model 

## Visualizing and Modeling Diamond prices 
Without detailed instruction, try to build a linear model of at least three variables to predict the worth of a diamond.


```{r load_diamonds_data, include = FALSE}
data(diamonds, package = "ggplot2")
names(diamonds)
```

Examine diamond price 
```{r summary_diamonds, echo = FALSE}
summary(diamonds)
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of diamond price") +
  xlab ('Diamond Price (in US Dollars)')
```

Examine diamond carat 


Examine diamond price and carat 


Examine the relationship between diamond price and cut 

